---
title: "2.cleaning_input_data"
author: "Emily Yeo"
date: "`r Sys.Date()`"
output: html_document
---

# Load packages

```{r}
#clear environment
rm(list = ls())
# load packages 
pacman::p_load(knitr, tidyverse, magrittr, lme4, lmerTest, GGally, corrplot, 
               Hmisc, kableExtra, dplyr, plyr, janitor, lubridate, survminer, 
               ggplot2, here, readr, readxl)
```

# Table 1 (main demographics)

```{r warning=FALSE}
#set to files directory
setwd("/Volumes/IPHY/ADORLab/Lab\ Projects/CHLAneuro/CHLAclean/rawdata")
# Table 1 Unique Patient info ####
Table_1 <- read.csv("Table1_UniquePatients_0615.csv",                     
                    na = c("NA", "don't know", "","    .",
                           "Missing: Not provided",999.00)) %>%
            dplyr::rename(Research_ID = "Research.ID",
                          DOB = "Date.of.Birth",
                          Sex = "Gender",
                          Race = "Race.At.First.Encounter",
                          Ethnicity = "Ethnicity.At.First.Encounter",
                          Language = "Preferred.Language.At.First.Encounter",
                          DateOfFirstHRIFVisit = "Date.Of.First.HRIF.Visit",
                          DateOfLastHRIFVisit = "Date.Of.Last.HRIF.Visit",
                          Gestational_Age1 = "Gestational.Age",
                          Gestational_Age_Date = "Gestational.Age.Date",
                          first_bayleys_age ="Age.at.First.Bayley.in.Months")

# Variable selection
Table_1 <- Table_1[ , c("Research_ID","DOB","Sex","Race","Ethnicity",
                        "Language",
                        "first_bayleys_age",
                        "DateOfFirstHRIFVisit",
                        "DateOfLastHRIFVisit", 
                        "Gestational_Age1")]
# Total Infants: 1561
length(unique(Table_1$Research_ID))
```

### Recode race and ethnicity 
This is shown in supplementary table 1. 

```{r}
Table_1 <- Table_1 %>% 
  mutate(V3_RACE = case_when(Ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                             Ethnicity == "Hispanic" ~ "Hispanic or Latino",
                             Ethnicity == "Latino" ~ "Hispanic or Latino",
                             Race == "Mexican/Mexican-American, Chican(x)" ~ "Hispanic or Latino",
                             Race == "Other Latin(x) and/or Hispanic" ~ "Hispanic or Latino",
                             Race == "Latino/a/x and/or Hispanic" ~ "Hispanic or Latino",
                             Race == "Chinese" ~ "Asian or Pacific Islander",
                             Race == "Other Asian" ~ "Asian or Pacific Islander",
                             Race == "Native Hawaiian" ~ "Asian or Pacific Islander",
                             Race == "Filipino" ~ "Asian or Pacific Islander",
                             Race == "Asian/Pacific" ~ "Asian or Pacific Islander",
                             Race == "Asian" ~ "Asian or Pacific Islander",
                             Race == "Native American" ~ "American Indian or Alaska Native",
                             Race == "African" ~ "Black or African American",
                             Race == "Other Black and/or African American" ~ "Black or African American",
                             Race == "Black and/or African American" ~ "Black or African American",
                             Race == "Middle Eastern and/or North African"	~ "Middle Eastern or North African",
                             Race == "Armenian" ~ "White",
                             Race == "White" ~ "White",
                             Race == "Other White" ~ "White",
                             Race == "Other" ~ "Other Race",
                             Race == "Declines to State" ~ "Nonconforming data",
                             Race == "None Specified" ~ "Nonconforming data",
                             Race == "Unknown" ~ "Nonconforming data"))

table(Table_1$V3_RACE)
# how many "other race" and hispanic ethnicity
filtered_data <- subset(Table_1, Race == "Other" & Ethnicity == "Hispanic or Latino")
count_rows <- nrow(filtered_data)
print(count_rows) # No instances of Race being Other & Ethnicity being Hispanic or Latino 
```

# Gestational Age Data
This was given later, in addition to the original meta data gestational ages. 

```{r}
#set to files directory
setwd("/Volumes/IPHY/ADORLab/Lab\ Projects/CHLAneuro/CHLAclean/rawdata")
GA <- readxl::read_xlsx("HRIF_Study_GA.xlsx") %>% 
  dplyr::rename(Research_ID = "Research ID",
                Gestational_Age2 = "Gestational Age",
                GA_Days = "GEST_AGE_AT_BIRTH")

# Subset cols of interest
GA_1 <- GA[ , c("Research_ID", "Gestational_Age2", "GA_Days")]
```

### Merge Gestational Data to Main Meta 

```{r warning=FALSE}
# Merge new gestational ages to Table 1: 1663
T1_GA <- merge(Table_1, GA_1, 
            by = "Research_ID",
            all = TRUE)
length(unique(T1_GA$Research_ID)) # 1563

# Extra people from new GA dataset
setdiff(T1_GA$Research_ID, Table_1$Research_ID) # "R14815068", "R59519385"

# just reorddering columns 
T1_GA <- T1_GA %>%
  dplyr::select(Research_ID, 
                Gestational_Age1, 
                Gestational_Age2,
                GA_Days,
                everything())

# Convert GA_days to weeks
T1_GA$GA_Days
T1_GA <- T1_GA %>% 
  mutate(GA_weeks = ifelse(!is.na(GA_Days) & GA_Days != 0, 
                           round(GA_Days / 7, 1), NA))

T1_GA$GA_weeks

# Create a new column "GA_combined" by merging "Gestational_Age2" into "Gestational_Age1"
T1_GA$GA_combined <- ifelse(is.na(T1_GA$Gestational_Age1), 
                         T1_GA$Gestational_Age2, 
                         T1_GA$Gestational_Age1)

T1_GA$GA_combined2 <- ifelse(is.na(T1_GA$GA_combined), 
                         T1_GA$GA_weeks, 
                         T1_GA$GA_combined)

# just reorddering columns 
T1_GA_GA<- T1_GA %>%
  dplyr::select(Research_ID, 
                Gestational_Age1, 
                Gestational_Age2,
                GA_weeks,
                GA_combined,
                GA_combined2) %>% 
        distinct()

summary(T1_GA_GA[c("Gestational_Age1", 
                "Gestational_Age2",
                "GA_weeks",
                "GA_combined",
                "GA_combined2")])

# Drop the original GAs we no longer need
Table_1 <- Table_1 %>% 
           dplyr::select(-c("Gestational_Age1"))

# Merge in new combined GAs ad meta data 
Table_1 <- merge(Table_1, T1_GA_GA, 
                 by = "Research_ID",
                 all = TRUE) 

# Remove the 2 new IDs that have no meta data
Table_1 <- Table_1[!(Table_1$Research_ID %in% 
                       c("R14815068", "R59519385")), ]

# How many children are missing GA scores? 89
sum(is.na(Table_1$GA_combined2)) #89
sum(!is.na(Table_1$GA_combined2)) #1472
length(unique(Table_1$Research_ID)) #1561

# How many children have GA scores, which are less than 20? 4
low_GAs <- subset(Table_1, GA_combined2 <= 20)

# Mention this in the manuscript
Table_1_above20 <- Table_1[Table_1$GA_combined2 >= 20 | is.na(Table_1$GA_combined2), ]
length(unique(Table_1$Research_ID)) # missing GA and 4 invalid 
```

# Participant Location

```{r warning=FALSE}
#set to files directory
setwd("/Volumes/IPHY/ADORLab/Lab\ Projects/CHLAneuro/CHLAclean/rawdata")
# Table 2 Unique Patient Zipcodes ####
Table_2 <- read.csv("Table2_UniqueHRIFVisit_0615.csv",               
                    na = c("NA", "don't know", "","    .",
                           "Missing: Not provided",999.00)) %>%
  dplyr::rename(Research_ID = "Research.ID",
                Encounter_ID = "Encounter.ID",
                DOB = "Date.of.Birth",
                Insurance = "Insurance.Type",
                Encounter_Age_Months = "Age.at.Encounter.in.Months",
                Encounter_Date = "Encounter.Date") %>% 
  mutate(Encounter_ID = as.character(Encounter_ID)) %>% 
  mutate(Encounter_Date = as.Date(Encounter_Date, format = "%m/%d/%Y"))

# Variable selection
Table_2 <- Table_2[ , c("Research_ID","Zipcode","Encounter_ID",
                        "Encounter_Age_Months","Encounter_Date")]

# Drop duplicate rows
Table_2 <- Table_2[!duplicated(Table_2), ]

# Drop rows with NA for all columns
Table_2 <- Table_2[rowSums(is.na(Table_2)) != ncol(Table_2), ]

# Create a column indicating ordinal rank of zipcode
Table_2 <- Table_2 %>%
  group_by(Research_ID) %>%
  mutate(ZipcodeRank = rank(Encounter_Date))
```

# First Encounters

```{r warning=FALSE}
#set to files directory
setwd("/Volumes/IPHY/ADORLab/Lab\ Projects/CHLAneuro/CHLAclean/rawdata")
Table_3 <- read_csv("Table3_DiagnosesFirstHRIF_0615.csv",               
                    na = c("NA", "don't know", "","    .",
                           "Missing: Not provided",999.00)) %>%
          dplyr::rename(Research_ID = `Research ID`,
                        Encounter_ID = `Encounter ID`,
                        Dx_Code = `Dx Code`,
                        Dx_Display = `Dx Display`)%>% 
          mutate(Encounter_ID = as.character(Encounter_ID))
```

# All encounters 

```{r warning=FALSE}
#Table 4 Last Visits ####
#set to files directory
setwd("/Volumes/IPHY/ADORLab/Lab\ Projects/CHLAneuro/CHLAclean/rawdata")
Table_4 <- read_csv("Table4_DiagnosesLastHRIF_0615.csv",                 
                    na = c("NA", "don't know", "", " .", 
                           "Missing: Not provided",999.00)) %>%
  dplyr::rename(Encounter_Age = `Age at Encounter in Months.`,
                Research_ID = `Research ID`,
                Encounter_ID = `Encounter ID`,
                Encounter_Date = `Encounter Date`,
                DOB = `Date of Birth`,
                Dx_Code = `Dx Code`,
                Dx_Display = `Dx Display`) %>% 
  mutate(Encounter_ID = as.character(Encounter_ID))
```

# Bayleys Scores of Cognitive Developement

```{r warning=FALSE}
#set to files directory
setwd("/Volumes/IPHY/ADORLab/Lab\ Projects/CHLAneuro/CHLAclean/rawdata")
Table_5 <- read_csv("Table5_BayleyScores_0615.csv",                 
                    na = c("NA", "don't know", "", " .", 
                           "Missing: Not provided",999.00)) %>%
  dplyr::rename(Research_ID = `Research ID`,        
                Gestational_Age = `Gestational Age.`,
                Baseline_Age = `Age at Bayley Administration in Months`,
                Birth_Date = `Date of Birth`,
                Bayley_Date = `Bayley Administration Date`,
                Motor_Composite_Score = `Motor Composite Score`,
                Lang_Composite_Score = `Language Composite Score`,
                Cognitive_Composite_Score = `Cognitive Composite Score`,
                Encounter_ID = `Encounter ID`) %>%
  mutate(Encounter_ID = as.character(Encounter_ID),
         FIN = as.character(FIN),
         Motor_Composite_Score = as.numeric(Motor_Composite_Score),
         Lang_Composite_Score = as.numeric(Lang_Composite_Score),
         Cognitive_Composite_Score = as.numeric(Cognitive_Composite_Score))
# Ignore warnings for irrelevant variables

# Select the variables of interest
Table_5 <- subset(Table_5, select = c("Research_ID",
                                      "Baseline_Age","Birth_Date","Encounter_ID",
                                      "Bayley_Date","Motor_Composite_Score", 
                                      "Lang_Composite_Score",
                                      "Cognitive_Composite_Score"))
length(unique(Table_5$Research_ID)) # 792

#format dates 
Table_5$Birth_Date = mdy(Table_5$Birth_Date)
Table_5$Bayley_Date = mdy(Table_5$Bayley_Date)

# compute the Bayley's age in Days
Table_5$Bayleys_Age_Days = as.numeric(Table_5$Bayley_Date - Table_5$Birth_Date)

# compute the Bayley's Age in Months 
Table_5$Bayleys_Age_Months <- interval(Table_5$Birth_Date, 
                                       Table_5$Bayley_Date) %/% months(1)

#removing participants older than 42 months at Bayleys test: 
sum(Table_5$Bayleys_Age_Months >= 42, na.rm = T) # Removed 11 over the age limit
Table_5 <- Table_5[Table_5$Bayleys_Age_Months <= 42,]
length(unique(Table_5$Research_ID)) # 792 (stays the same because of multiple visits for same IDs)

#	How many NAs per Bayley's Test
sum(is.na(Table_5$Motor_Composite_Score)) #348
sum(is.na(Table_5$Lang_Composite_Score)) #369
sum(is.na(Table_5$Cognitive_Composite_Score)) #567

# Convert Bayleys to integer 
Table_5$Motor_Composite_Score <- as.integer(Table_5$Motor_Composite_Score)
Table_5$Lang_Composite_Score <- as.integer(Table_5$Lang_Composite_Score)
Table_5$Cognitive_Composite_Score <- as.integer(Table_5$Cognitive_Composite_Score)

sum(is.na(Table_5$Motor_Composite_Score)) #348
sum(is.na(Table_5$Lang_Composite_Score)) #369
sum(is.na(Table_5$Cognitive_Composite_Score)) #567

# Subset any absurdly high scores
over_150 <- subset(Table_5, Lang_Composite_Score > 150 | 
                     Motor_Composite_Score > 150 | 
                     Cognitive_Composite_Score > 150)
nrow(over_150) #2

# remove kids with scores out of range 
Table_5 <- subset(Table_5, Lang_Composite_Score <= 150 | 
                    is.na(Table_5$Lang_Composite_Score)) 

Table_5 <- subset(Table_5, Motor_Composite_Score <= 150 | 
                       is.na(Table_5$Motor_Composite_Score))

Table_5 <- subset(Table_5, Cognitive_Composite_Score <= 150 | 
                    is.na(Table_5$Cognitive_Composite_Score))

length(unique(Table_5$Research_ID)) # 791

# Add column indicating date rank (only 1 missing rank)
Table_5 <- Table_5 %>%
  group_by(Research_ID) %>%
  mutate(DateRank = rank(Bayley_Date))

# "R77029329" this person has 2 dates and 5 rows. ranks ar 2 and 4.5 
random_dude <- subset(Table_5, 
                      Research_ID == "R77029329" & Lang_Composite_Score == 77)

# Subset first bayley's visit only
Table_5 <- subset(Table_5, DateRank == 1|
                    DateRank == 1.5) 

# merge rando back into dataset
Table_5 <- rbind(Table_5, random_dude)

# Remove date rank column (unnecessary)
Table_5 <- Table_5[, !names(Table_5) %in% c("DateRank")]
```

# Child Opportunity Scores

```{r warning=FALSE}
#set to files directory
setwd("/Volumes/IPHY/ADORLab/Lab\ Projects/CHLAneuro/CHLAclean/rawdata")
Table_COI <- read.csv("Zip_Code_COI.csv",
                      na = c("NA", "don't know", "", "    .", 
                             "Missing: Not provided", 999.00)) %>%
            dplyr::rename(Zipcode = "ZIP.Code",
                          RC_Assigned = "RC.assingments",
                          RC1 = "Regional.Center.1",
                          RC2 = "Regional.Center.2",
                          RC3 = "Regional.Center.3",
                          asian = "aian") %>% 
  mutate(Zipcode = as.character(Zipcode))

# Variable selection
Table_COI <- Table_COI[ , c("Zipcode","r_coi_nat","c5_coi_nat")]
```

### Join COI and meta data 

```{r}
Table_all <- Table_1_above20 %>% 
  left_join(Table_2, by = "Research_ID") %>% 
  left_join(Table_COI, by = "Zipcode")

#length(unique(Table_all$Research_ID))
#sum(is.na(Table_all$r_coi_nat)) 
#sum(!is.na(Table_all$r_coi_nat)) 
```

# Load PMCA classifications
Classifications were determined using script 1

```{r}
pmca_first <- read_csv("/Users/emily/projects/research/ADOR/CHLA_clean/outputs/dataframes/pmca_first_12.05.csv", 
                       na = c("NA", "don't know", "", " .", 
                              "Missing: Not provided",999.00)) %>% 
                       dplyr::rename(pmca_first = "PMCA_category")

length(unique(pmca_first$Research_ID)) # 1549
```

### Merge PMCA data and meta

```{r}
Table_all2 <- Table_all %>% left_join(pmca_first)
sum(unique(Table_5$Research_ID) %in% pmca_first$Research_ID)
```

### Merge PMCA, COI, Bayleys and Meta

```{r}
T1_noEnc <- Table_all2 %>% dplyr::select(-c(Encounter_ID, Encounter_Date, ZipcodeRank))
T5_noEnc <- Table_5 %>% dplyr::select(-c(Encounter_ID))

T1_5 <- merge(T1_noEnc, T5_noEnc, 
              by = "Research_ID",
              all = TRUE) %>% 
              unique()
```

#### A few checks

```{r}
sum(is.na(T1_5$r_coi_nat)) # 389
sum(!is.na(T1_5$r_coi_nat)) # 1423
length(unique(T1_5$Research_ID)) # 1558

# How many not missing GA
subset_GA <- T1_5[!is.na(T1_5$GA_combined2), ]
length(unique(subset_GA$Research_ID)) # 1468

# Identifying how many infants have at least 1 Bayleys score
subset_bay <- Table_5[(!is.na(Table_5$Motor_Composite_Score) | 
                         !is.na  (Table_5$Lang_Composite_Score)  | 
                         !is.na(Table_5$Cognitive_Composite_Score)), ]

# Identifying how many infants have all 3 Bayleys scores
subset_bay_and <- Table_5[(!is.na(Table_5$Motor_Composite_Score) & 
                             !is.na(Table_5$Lang_Composite_Score)  & 
                             !is.na(Table_5$Cognitive_Composite_Score)), ]

```

### More edits to bayleys and meta 

```{r}
# Joining and editing a master table ####
table_df <- Table_all2
length(unique(table_df$Research_ID)) # 1557
bayleys <- Table_5

# Create a column indicating rank of bayleys ages 
bayleys <- bayleys%>%
  dplyr::group_by(Research_ID) %>%
  mutate(bayleys_age_rank = rank(Bayleys_Age_Days)) 

# Merge bayley's data and meta data
# This gets used later to assess ages at bayleys and encounters.
meta_for_testing <- table_df %>% 
  dplyr::select(Research_ID, Sex, Race, Ethnicity, V3_RACE, Language,  
                Gestational_Age1, Gestational_Age2, GA_combined,
                GA_combined2, r_coi_nat, pmca_first, Encounter_Age_Months,
                first_bayleys_age, Encounter_Date, Zipcode, DateOfFirstHRIFVisit) %>% 
  unique() %>% 
  filter(Research_ID %in% bayleys$Research_ID)

meta_for_bayleys <- table_df %>% 
  dplyr::select(Research_ID, Sex, Race, Ethnicity, V3_RACE, Language,  
                Gestational_Age1, Gestational_Age2, GA_combined,
                GA_combined2, r_coi_nat, pmca_first) %>% 
  unique() %>% 
  filter(Research_ID %in% bayleys$Research_ID)

meta_bayleys <- merge(meta_for_bayleys, bayleys, 
                      by.x = "Research_ID", 
                      by.y = "Research_ID")

```

### Making categories 
out of continuous variables 

```{r}
## reorder pmca cats for plotting later
meta_bayleys$pmca_first <- factor(meta_bayleys$pmca_first, 
                                  levels = c("Non-chronic", 
                                             "Non-complex Chronic", 
                                             "Complex Chronic"))

# Breaking up COI
quantile(meta_bayleys$r_coi_nat, c(0:3/3), na.rm=TRUE)
meta_bayleys$coi_cat <- ifelse(meta_bayleys$r_coi_nat <= 16,"low",
                               ifelse(meta_bayleys$r_coi_nat > 16 & meta_bayleys$r_coi_nat <= 41.4, "medium",
                                      ifelse(meta_bayleys$r_coi_nat > 41.4, "high", NA)))

### Gestational Age Categories ####
meta_bayleys$Gestational_Class = ifelse(test = meta_bayleys$GA_combined2 < 28,
                                        yes = "Extremely_Preterm",
                                        no = ifelse(test = meta_bayleys$GA_combined2 < 32,
                                                    yes = "Very_Preterm",
                                                    no = ifelse(test = meta_bayleys$GA_combined2 < 34,
                                                                yes = "Moderate_Preterm",
                                                                no = ifelse(test = meta_bayleys$GA_combined2 <= 37,
                                                                            yes = "Late_Preterm",
                                                                            no = "Term"))))    
## reorder cats for plotting later
meta_bayleys$Gestational_Class <- factor(meta_bayleys$Gestational_Class, 
                                         levels = c("Extremely_Preterm", 
                                                    "Very_Preterm",
                                                    "Moderate_Preterm",
                                                    "Late_Preterm",
                                                    "Term"))

# Create a numeric variable for PMCA ranked 
meta_bayleys$PMCA_rank <- ifelse(meta_bayleys$pmca_first=="Non-chronic",0,
                                 ifelse(meta_bayleys$pmca_first=="Non-complex Chronic",1,
                                        ifelse(meta_bayleys$pmca_first=="Complex Chronic",2, NA)))

```

# Subset to only first Bayleys Assessments 

```{r}
## Update meta_bayleys with ranks
meta_1_date_bayleys <- meta_bayleys %>%
  group_by(Research_ID) %>%
  mutate(
    bayleys_age_rank = rank(Bayleys_Age_Days, ties.method = "max"),
    DateRank = rank(Bayley_Date)
  ) %>%
  ungroup() 

## Define rank values and dataframe names for age ---
age_ranks <- 1:3
age_df_names <- paste("meta", age_ranks, "age_bayleys", sep = "_")

# Create age dataframes
lapply(seq_along(age_ranks), function(i) {
  assign(x = age_df_names[i], 
         value = filter(meta_1_date_bayleys, bayleys_age_rank == age_ranks[i]), 
         envir = .GlobalEnv)
})

## Define rank values and dataframe names for date ----
date_ranks <- 1:3
date_df_names <- paste("meta", date_ranks, "date_bayleys", sep = "_")

# Create date dataframes
lapply(seq_along(date_ranks), function(i) {
  assign(x = date_df_names[i], 
         value = filter(meta_1_date_bayleys, DateRank == date_ranks[i]), 
         envir = .GlobalEnv)
})
```

# Determine age at Bayleys

```{r}
# subset only the cols we want 
bayley_age <- meta_for_testing %>% 
  dplyr::select("Research_ID","first_bayleys_age") %>% 
  unique()

# This makes sure we only using the first bayleys assessment. 
meta1bay_age <- merge(meta_1_date_bayleys, bayley_age, by = "Research_ID",
                      all.x = TRUE, all.y = FALSE) %>% 
  dplyr::select(c("Research_ID","first_bayleys_age"))
```

# Determine age at first encounter

```{r}
T2 <- Table_2 %>% 
  dplyr::select(c("Research_ID", "Encounter_ID", 
                  "Encounter_Age_Months")) %>% unique()
T3 <- Table_3 %>% 
  dplyr::select(c("Research_ID", "Encounter_ID")) %>% unique()

T2_3 <- merge(T2, T3, 
              by = c("Encounter_ID", "Research_ID"), all.y = TRUE)

# merge bayles and encounter ages 
T_age <- merge(T2_3, meta1bay_age, by = c("Research_ID"), 
               all.x = TRUE, all.y = FALSE)
```

# Write main analysis files 

```{r warning=FALSE}
# Write tis data file with everything in it 
#setwd("/Volumes/IPHY/ADORLab/Lab\ Projects/CHLAneuro/CHLAclean/clean_inputs/july5/")
#save(meta_bayleys,
#     meta_1_date_bayleys,
#     meta_for_testing,
#     T_age,
#     file='meta_bay_checks_july5.RData')
```

# After Reviews

## 1. Make a descriptive table of those with vs. without Bayleys Scores

```{r}
# Filter Table_1_above20 to keep only rows with Research_ID not in Table_5
no_bayley <- table_df %>%
  anti_join(subset_bay_and, by = "Research_ID") %>% 
  mutate(has_bay = "no") %>% 
  anti_join(Table_5, by = "Research_ID")

has_bay <- table_df %>% 
  anti_join(no_bayley, by = "Research_ID") %>% 
  dplyr::mutate(has_bay = "yes")

both <- dplyr::bind_rows(no_bayley, has_bay)

# sum of these is 1556
unique(no_bayley$Research_ID) %>% length()
unique(has_bay$Research_ID) %>% length()
length(intersect(no_bayley$Research_I, has_bay$Research_ID))

## reorder pmca cats for plotting later
both$pmca_first <- factor(both$pmca_first, 
                                  levels = c("Non-chronic", 
                                             "Non-complex Chronic", 
                                             "Complex Chronic"))
# Breaking up COI
quantile(both$r_coi_nat, c(0:3/3), na.rm=TRUE)
both$coi_cat <- ifelse(both$r_coi_nat <= 16,"low",
                               ifelse(both$r_coi_nat > 16 & both$r_coi_nat <= 41.4, "medium",
                                      ifelse(both$r_coi_nat > 41.4, "high", NA)))

### Gestational Age Categories ####
both$Gestational_Class = ifelse(test = both$GA_combined2 < 28,
                                        yes = "Extremely_Preterm",
                                        no = ifelse(test = both$GA_combined2 < 32,
                                                    yes = "Very_Preterm",
                                                    no = ifelse(test = both$GA_combined2 < 34,
                                                                yes = "Moderate_Preterm",
                                                                no = ifelse(test = both$GA_combined2 <= 37,
                                                                            yes = "Late_Preterm",
                                                                            no = "Term"))))    
## reorder cats for plotting later
both$Gestational_Class <- factor(both$Gestational_Class, 
                                         levels = c("Extremely_Preterm", 
                                                    "Very_Preterm",
                                                    "Moderate_Preterm",
                                                    "Late_Preterm",
                                                    "Term"))


both_unique <- both %>% dplyr::select(Research_ID, Sex, GA_combined2,
                  V3_RACE, coi_cat, pmca_first, has_bay) %>% unique()

769 + 796
```

Make a table comparing those who do vs. dont have Bayleys scores 

```{r}
# which vars to include 
descrip_vars <- c("Sex", "GA_combined2",
                  "V3_RACE",
                  "coi_cat",
                  "pmca_first", 
                  "has_bay")
desc_labels <- c("Sex","Gestational Age",
                 "Race and/or Ethnicity",
                 "Child Opp Index",
                 "PMCA Category",
                 "Did they have Bayleys")

des_cat_vars <- c("Sex","V3_RACE","coi_cat","pmca_first", "has_bay")

descrip_table <- dplyr::select(both_unique, one_of(descrip_vars))

labelled::var_label(descrip_table) <- desc_labels
descriptive_T1 <- CreateTableOne(vars = descrip_vars, 
                                   data = descrip_table, 
                                   factorVars = des_cat_vars, 
                                   strata = "has_bay",
                                   includeNA=T,
                                   addOverall = T,
                                   test = TRUE,
                                   testApprox = chisq.test,
                                   testNormal = oneway.test,
                                   argsNormal = list(var.equal = TRUE),
                                   testNonNormal = kruskal.test)
#kable(summary(descriptive_T1))
d_T1 <- print(descriptive_T1,
               catDigits = 1, 
               contDigits = 1, 
               varLabels = T,
               exact = "stage", 
               quote = FALSE, 
               noSpaces = TRUE, printToggle = FALSE)
## Save to html
d_T1 <- d_T1 %>% 
     kable(caption = "Baseline Characteristics") %>%
     kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

d_T1
```




